---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sep

The goal of `sep` is to help collaborators and interested users to create graphics as used in the Swiss Environmental Panel Reports.

## Installation

You can install a development version of `sep` from [GitHub](https://github.com/) with:

```{r, message=FALSE}
# install.packages("devtools")
devtools::install_github("bonschorno/sep")
```

```{r}
library(sep)
library(tidyverse)
```


## A quick overview

Probably the most useful function is `sep_palette` for the time being. It is is a non-exhaustive selection of official color palettes as listed on the ETH Zurich website. This makes adding the official ETH colors to the plots very easy with the various `scale_*_manual` functions. 

```{r}
sep_palette("Darkblues")
```

```{r}
sep_palette("Pinks", n = 4)
```

```{r, fig.retina=4}
ggplot(mtcars, aes(x = disp, y = mpg, color = as.factor(carb))) +
  geom_point() +
  scale_color_manual(values = sep_palette("Standard"))
```

To make the plots not only more visually appealing, but also to standardize them across the various reports, `theme_sep` was designed. It can either be applied locally or set as default at the beginning of the script. 

```{r, fig.retina=4}
ggplot(mtcars, aes(x = disp, y = mpg, color = as.factor(carb))) +
  geom_point() +
  scale_color_manual(values = sep_palette("Standard")) +
  theme_sep()

#set as default theme
theme_set(theme_sep())
```

Two other functions are currently still very basic, but will be improved soon. One is `crosstable`, which creates simple crosstables of n (numeric) variables and `simple_barplot` to make creating bar plots very easy.

```{r}
crosstable(vars = c("gear", "am"), df = mtcars)
```


```{r example, fig.retina=4}
simple_barplot(df = mtcars, var = gear, title = "Gears", width = 0.3) 
```

