---
title: "Walking you through sep"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Walking you through sep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.retina = 4,
  fig.height = 6, 
  fig.width = 8
)
```

`sep` can be downloaded as follows:

```{r, eval=FALSE}
devtools::install_github("bonschorno/sep")
```

```{r setup}
library(sep)
library(dplyr)
library(ggplot2)
```

Currently, `sep` offers the following functions:

```{r, echo=FALSE}
ls("package:sep")
```

The prefixes indicate the step in the workflow. The functions beginning with `cb` simplify creating the codebook. The `eda` functions are for exploratory data analysis. The functions starting with `plot`, in turn, help to visualise the data. Finally, there are some additional helper functions.

FYI: The dataset used here is not publicly accessible for data protection reasons. However, it can be requested [here](https://istp.ethz.ch/research/sep.html).

```{r}
load("upanel_w5.RData")
```

Very often we need to know the relative frequencies of the responses to a survey item. `eda_freqtable` helps to get these numbers quickly. By default, the logical operator is `rm.dk = FALSE`, i.e. the relative proportion of "Don't know" responses is included.

```{r}
eda_freqtable(w5, item = w5_q27)
eda_freqtable(w5, item = w5_q27, rm.dk = T)
```

To calculate the relative frequencies based on two variables, you can use `eda_crosstable`. The two variables must be specified as characters in a vector. Here, too, "Don't knows" can be included/excluded via the logical operator `rm.dk`.

```{r}
eda_crosstable(w5, items = c("w5_q2", "w5_q11"))
eda_crosstable(w5, items = c("w5_q2", "w5_q11"), rm.dk = TRUE)
```

## Visualizations

Four plot functions are currently available: `plot_bar_v`, `plot_bar_h`, `plot_groupbar_v`, `plot_groupbar_h.` As the respective names imply, the first two functions are used to visualise one variable, while the last two allow visualisation with a grouping variable.

In their simplest forms, `plot_bar_v` and `plot_bar_h` plot the relative frequency of the item in a vertical and horizontal plot, respectively. The values of the items are not yet labelled and the proportion "Don't know" responses must also be specified.

```{r}
plot_bar_v(w5, item = w5_q11, textsize = 11)
```
 
There is a corresponding argument in the function for the share of "Don't know" responses (`dk_share`). For the labels, you can use the underlying `ggplot2` structure of the functions by simply adding further layers (in this case, `scale_x_discrete` to add labels). **NB: When using the horizontal bar plots (with one or two variables), you have to reverse the order of your manual labels to display them correctly!!** With further optional arguments, the bars' colour or the relative frequencies' position can be changed, among other things.

```{r}
plot_bar_h(w5, item = w5_q11) +
    scale_x_discrete(labels = rev(c("Item One", "Item Two", "Item Three", "Item Four", "Item Five")))
```

Before moving on to the visualisations of grouped variables, it's worth highlighting the helper function `sep_palette". This function allows users to choose from a range of official ETH colour palettes. Available are the following palettes:

`"Standard", "Darkblues", "Greens", "Lightblues", "Lightgreens", "Pinks"`

You can select the palette of your choice by either saving it as an object or directly calling it in the plot function.

```{r}
plot_groupbar_v(w5, item = w5_q11, by = w5_q2, labels = c("Female", "Male"), values = sep_palette("Standard")) +
 scale_x_discrete(labels = c("Item One", "Item Two", "Item Three", "Item Four", "Item Five"))
```

**Again, note that with the horizontal plots, you have to change reverse the order of the manual labels to display them correctly!**

```{r}
plot_groupbar_h(w5, item = w5_q11, by = w5_q2, labels = c("Female", "Male"), values = sep_palette("Darkblues")) +
 scale_x_discrete(labels = rev(c("Item One", "Item Two", "Item Three", "Item Four", "Item Five")))
```
