---
title: "Walking you through"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Walking you through}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

For the time being, the sep-package remains a private project, which is why it can only be installed via the following Github repository. 

```{r}
# install.packages("devtools") 
#devtools::install_github("bonschorno/sep")
```


```{r setup, include=FALSE}
library(sep)
```

The package is divided into different helper functions. Functions that start with `eda_*` help with exploratory data analysis. As an (optional) first step, the survey is loaded into the global environment. This is not good coding practice, but it has a significant advantage. This way, you do not always have to specify the survey in the data argument later, but only the corresponding survey item, which simplifies the workflow considerably. 

It must be added that the data set used is not publicly accessible for data protection reasons. However, it can be requested here.

```{r}
load("upanel_w5.RData")
eda_loadtoenv(w5)
```

We almost always need to know the relative frequencies of the responses to a survey item. `eda_freqtable` helps to get these numbers quickly. By default, the logical operator is `rm.dk = FALSE`, i.e. the relative proportion of "Don't know" responses is included.

```{r}
eda_freqtable(item = w5_q27)
eda_freqtable(item = w5_q27, rm.dk = T)
```

